---
title: "SQL"
output: html_document
---



##### On this page you will find examples on how to use SQL in R Studio to find specific information from a large database. For these examples, we will use the Lahman's Baseball Database.

***

## Filtering Rows and Columns 

With SQL, you can filter columns with SELECT and rows with WHERE. Let's look at an example using the Lahman Baseball Database. First, we need to load some packages.

```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```

Suppose we want to see the homerun totals for the 1927 Yankees. We could write the following:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```


Suppose we wanted to find all instances where Yankee players have hit 40 homeruns or more:
  

```{r}
query<- 'SELECT playerID,yearID,teamID,HR FROM Batting
WHERE TeamID="NYA" and HR>=40'
sqldf(query)
```

Suppose we want to find all players who have more than 40 home runs, and less than 60 strikeouts:
  

```{r}
query<- "SELECT playerID,yearID,teamID,HR,SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```

Suppose we want to find all players on the Phillies who have hit more than 30 homeruns in the 1970s:
  

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and HR>30 and yearID<=1979 and yearID>=1970"
sqldf(query)
```


***

## Ordering Rows

Now, let's add something new... Suppose we want to find all the players who have hit more than 50 homeruns, and we want the list to show the player with the highest number of homeruns first. The ORDER BY commands automatically orders in ascending order, so we have to add DESC for the highest number to appear first: 

```{r}
query<- "SELECT playerID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

Finally, suppose we want to find all the players who have been at bat more than 400 time, and have less than 10 strikeouts. We want to order the list by least to most number of strikeouts.

```{r}
query<- "SELECT playerID,yearID,teamID,HR,AB,SO FROM Batting
WHERE SO<10 and AB>400
ORDER BY SO"
sqldf(query)
```

***

## Aggregation 

Let's do some more examples using Aggregation as a way to group information. Supppose we want to find the total of all Babe Ruth's homeruns. We will use the command GROUP BY to find this information. The following commands will also be used thoughout these examples: sum(), avg(), max(), and min().

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"
sqldf(query)
```

Suppose we need to find the career homerun totals of all players, but limit the display to only those that have 600 or more homeruns. We want to have the players with the highest totals at the top. We will use the command HAVING to limit the display. 

```{r}
query<-"SELECT playerID,sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>600
ORDER BY sum(HR) DESC"
sqldf(query)
```


Finally, suppose we want to find the players with the highest homerun average over their careers. We want to limit the display to to those who have an average of more than 30 homeruns per year. Again, we want to see the players with the highest averages on top of the list. 

```{r}
query<-"SELECT playerID,avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```

***

## Joins

Joining is a way take information from two differnt tables and display them similanteously. So far, we've only been using the Batting table, but for these examples we will need to join the Batting table with the Master table, and the Teams table. We will use the command INNER JOIN to connect these tables, and ON to specifiy the common link between the two tables. For this first example, the common link is player ID, and we will use nameFIRST and nameLAST from the Master table. 


Suppose we want to see the home runs, year, and team ID, for Babe Ruth, but instead of displaying Babe Ruth's player ID, we want our results to display his first and last name.


```{r}
query<-"SELECT nameFIRST,nameLAST,teamID,yearID,HR
FROM Batting INNER JOIN Master
On Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"
sqldf(query)
```

For the last example, suppose we want to find all instances of players hitting more than 50 homeruns. We want to display the first and last names of players, the team ID, the year ID, and the homerun totals.

```{r}
query<-"SELECT nameFiRST,nameLAST,teamID,HR
FROM Batting INNER JOIN Master
On Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

Suppose we want to list all the players named Bob who averaged more than one million dollars per year in salary for their careers. Include last names. 

```{r}
query<-"SELECT nameLAST,nameFIRST,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFIRST='Bob'
GROUP BY Salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"

sqldf(query)
```

