---
title: "SQL"
output: html_document
---



##### On this page you will find examples on how to use SQL in R Studio to find specific information from a large database. For these examples, we will use the Lahman's Baseball Database.

***

## Filtering Rows and Columns 

With SQL, you can filter columns with SELECT and rows with WHERE. Let's look at an example using the Lahman Baseball Database. First, we need to load some packages.

```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```

Suppose we want to see the homerun totals for the 1927 Yankees. We could write the following:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```


Suppose we wanted to find all instances where Yankee players have hit 40 homeruns or more:
  

```{r}
query<- 'SELECT playerID,yearID,teamID,HR FROM Batting
WHERE TeamID="NYA" and HR>=40'
sqldf(query)
```

Suppose we want to find all players who have more than 40 home runs, and less than 60 strikeouts:
  

```{r}
query<- "SELECT playerID,yearID,teamID,HR,SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```

Suppose we want to find all players on the Phillies who have hit more than 30 homeruns in the 1970s:
  

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and HR>30 and yearID<=1979 and yearID>=1970"
sqldf(query)
```


***

## Ordering Rows

Now, let's add something new... Suppose we want to find all the players who have hit more than 50 homeruns, and we want the list to show the player with the highest number of homeruns first. The ORDER BY commands automatically orders in ascending order, so we have to add DESC for the highest number to appear first: 

```{r}
query<- "SELECT playerID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"
sqldf(query)
```

Finally, suppose we want to find all the players who have been at bat more than 400 time, and have less than 10 strikeouts. We want to order the list by least to most number of strikeouts.

```{r}
query<- "SELECT playerID,yearID,teamID,HR,AB,SO FROM Batting
WHERE SO<10 and AB>400
ORDER BY SO"
sqldf(query)
```

***

## Aggregation 


